A unified model for parsing and sentence understanding

### 2016-03-14 writing/hard_stack_paper/paper.tex
First stab at thin-stack explanation

%%% Custom additions %%%
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{ifthen}
\usepackage{framed}

@@

\paragraph{Representing the stack in memory} Or, thin stack. And the cursor strategy for the buffer. 
\todo{[JG] Explain.}

\vspace{8em}

@@

>>>>

%%% Custom additions %%%
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{ifthen}
\usepackage{framed}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

@@

@@

\subsection{Efficient stack representation}

A naive implementation of the PINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift and \reduce operations only affect the top of the stack).

We designed an alternative minimal-space stack algorithm inspired by the zipper technique \cite{huet1997zipper}. For a single input sentence, we represent the stack over the entire feedforward with a $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

\begin{algorithm}
\caption{The thin-stack algorithm}
\label{alg:thin-stack}
\begin{algorithmic}[1]
  \Function{Step}{bufferTop, op, $t$, $S$, $Q$}
    \If{op = \shift}
      \State $S$[$t$] := bufferTop
    \ElsIf{op = \reduce}
      \State right := $S$[$Q$.pop()]
      \State left := $S$[$Q$.pop()]
      \State $S$[$t$] := \Call{Compose}{left, right}
    \EndIf
    \State $Q$.push($t$)
  \EndFunction
\end{algorithmic}
\end{algorithm}

\begin{table}
\centering
\begin{tabular}{c|cl}
  \toprule
  $t$ & $S$[$t$] & $Q_t$ \\
  \midrule
  0 & $a$ & 0 \\
  1 & $b$ & 0 1 \\
  2 & $c$ & 0 1 2 \\
  3 & $M(c, b)$ & 0 3 \\
  4 & $M(M(c, b), a)$ & 4 \\
  \bottomrule
\end{tabular}
\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce sequence on the input sentence $(a, b, c)$. $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$. The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
\label{tbl:thin-stack}
\end{table}
